---
title: "STAT 631 Homework 7"
author: "Jack Cunningham (jgavc@tamu.edu)"
date: 10/21/2024
date-format: short
format:
  pdf:
    include-in-header:
      - text: |
          \usepackage{amsmath}
editor: visual
engine: knitr
---

1\)

```{r}
load("HW07.RData")
n = dim(Rt)[1];N = dim(Rt)[2]
```

The first thing we must do is create our vectors for $Y_{jt}=R_{jt}-\mu_{ft}$, the excess returns on the jth security. And $Y_{Mt}=R_{Mt}-\mu_{ft}$ the excess returns on the market portfolio.

```{r}
Yt = apply(Rt, 2, function(u) u - Rf)
YM = RM - Rf
```

Next we will compute the linear model:

$$
Y_{t}=\alpha+\beta(Y_{Mt})+e_t
$$

```{r}
fit = lm(Yt ~ YM)
sfit = summary(fit)
```

a\)

The estimates of beta for each asset is below:

```{r}
beta_est <- fit$coefficients[2,]
beta_est
```

The assets with the highest betas are Micron Technology, Inc. and Polaris Industries Inc. with $\hat{\beta}=$ `r c(beta_est[6], beta_est[8])` respectively.

b\)

The square risk is computed through the below formula:

$$
\sigma_{j}^2=\beta_j^2\sigma_M^2+\sigma_{e_j}^2
$$\
Which has two components, market of systematic component of risk $\beta_j^2 \sigma_M^2$ and unsystematic component of risk $\sigma^2_{e_j}$. In the regression context the proportion of square risk that is due to the systematic risk is equivalent to $R^2$ for each regression.

```{r}
R.Sq = c()
for(i in 1:N){
  R.Sq = c(R.Sq, sfit[[i]]$r.sq)
}
names(R.Sq) = syb; R.Sq
```

c\)

If we assume that each $Y_{it}$ follows CAPM that means we are assuming that $\alpha_i=0$ for each. Then the estimate of excess return is simply $\hat{Y_{it}}=\hat{\beta_i}\bar{Y}_{Mt}$.

```{r}
average_market_return <- mean(YM)
excess_return <- beta_est*average_market_return
excess_return
```

d\)

To test if the CAPM holds for each individual asset we test:

$$
H_0:\alpha_i=0,H_1:\alpha_i\neq0
$$

This is automatically computed through the lm function.

```{r}
Alpha = c()
for(i in 1:N){
  Alpha = rbind(Alpha, sfit[[i]]$coef[1,])
}
rownames(Alpha) = syb; Alpha
```

At a significance level of $\alpha = 0.05$ we can say that every asset except Broadcom Inc. cannot reject the null hypothesis. For Broadcom Inc the estimate of $\alpha$ is `r Alpha[2,1]` . Since this is greater than zero we can conclude that this security was under-priced in the past.

2\)

a\)

If we assume $y_1,…,y_m$ are i.i.d k x 1 random vectors and $y_i\sim N_k(0,\Omega)$ Then $S=\sum_{i=1}^my_iy_i^T \sim W_k(m,\Omega)$, the Wishart distribution with N degrees of freedom. Then:

$$
\frac{m-k+1}{k}y^TS^{-1}y \sim F_{k,m-k+1}
$$

We are testing:

$$
H_0:\alpha=0,H_1:\text{Any } \alpha_i\neq0
$$

For the Wald test we have the below:

$$
F_W=\frac{n-N-1}{nN}\frac{1}{m_{11}}\hat{\alpha}^T\hat{\Sigma_{e}}^{-1}\hat{\alpha}\sim F_{N,n-N-1}
$$

```{r}
alpha = Alpha[, "Estimate"]
et = resid(fit)
Sig = 1/n*t(et)%*%et
m11 = sfit[[1]]$cov.unscaled[1,1]
wald = (n - N - 1)/(n*N)*1/m11*t(alpha)%*%solve(Sig)%*%alpha
cat("Wald test:");c(statistic = wald, p.value = 1 - pf(wald, N, n - N - 1))
```

The likelihood ratio test statistic is the discrepancy between the full and reduced model under $H_0$:

$$
\text{LRT}=2\{\ell(\hat{\alpha},\hat{\beta},\hat{\Sigma_e})-\ell(0,\hat{\beta_0},\hat{\Sigma}_{0e})\}
$$

Through simplification and a non-normality adjustment we have:

$$
\Im_{LR}=(n-N/2-2)\{\log|\hat{\Sigma}_{0e}|-\log|\hat{\Sigma}_e|\} \sim X_N^2
$$

```{r}
fit0 <- lm(Yt ~ YM - 1)
et0 <- resid(fit0)
Sig0 = 1/n*t(et0)%*%et0
lr = (n-N/2 - 2)*(log(det(Sig0))-log(det(Sig)))
cat("Likelihood ratio test:"); c(statistic = lr, p.value = 1 - pchisq(lr,N))
```

Both tests fail the reject the null hypothesis that the CAPM model holds at significance level $\alpha = 0.05$. This is different than what we would expect as we previously found that individually Broadcom Inc. did not follow CAPM. However these test the overall p-value as opposed to the individual p-values from earlier.

b\)

We have been unable to reject the null hypothesis that the CAPM model holds. In this case the total risk of $Y_t$ is:

$$
\sigma_{j}=\sqrt{\beta_j^2\sigma_M^2+\sigma_{e_j}^2}
$$

We can find the systematic component of risk through$\sqrt{\beta^2_j\sigma^2_M}$, along with the covariance: $\sigma_{jj^\prime}=\beta_j\beta_{j^\prime}\sigma^2_M$:

```{r}
systematic <- beta_est%*%t(beta_est)*as.numeric(var(YM))
systematic
```

The non-systematic component of risk is $\sigma^2_{e_j}$, this is the variance of errors $\hat{\Sigma}$:

```{r}
Sig
```

c)

We can find the minimum non-systematic variance portfolio by using $\Sigma$ from our regression in the minimum variance portfolio formula.

$$
w_{min.v}=\frac{\Sigma^{-1}1}{1^T\Sigma^{-1}1}
$$

```{r}
ones = rep(1, N)
IS = solve(Sig)
a = as.numeric((t(ones)%*%IS%*%ones))
w.min = 1/a*(IS%*%ones)
w.min
```

d\)

```{r}
w_equal <- rep(1/8,N)
```

To test both portfolios we create 8 x 2 matrix W:

```{r}
W = cbind(w.min,w_equal)
```

Then the hypothesis to test if these two portfolios follow the CAPM model is:

$$
H_0:W^T\alpha=0,H_A:W^T\alpha \neq0
$$

Then we have the test statistic:

$$
\frac{m-k+1}{k}y^TS^{-1}y \sim F_{k,m-k+1}
$$

Where $y=W^T\hat{\alpha},S=n\widehat{var}(W^T \hat{\alpha}),m=n-2$. With $var(\hat{\alpha})=m_{11}\Sigma_e$. And $m_{11}$ is $(X^TX)^{-1}_{11}$.

```{r}
y = t(W)%*%alpha
S = n*m11*t(W)%*%Sig%*%W
m = n - 2
k = 2

test_stat <- (m-k+1)/k*t(y)%*%solve(S)%*%y
cat("Wald test statistic: "); c(statistic = test_stat, 
                               p.value = 1 - pf(test_stat, k, n - k - 1))
```

We cannot reject the null hypothesis that both portfolios follow CAPM.

3\)

```{r}
load("Midterm22.RData")
```

a\)

```{r}
Yt = apply(rt, 2, function(u) u - Rf)
YM = Mkt - Rf
fit = lm(Yt ~ YM)
betas = fit$coefficients[2,]
ind = order(betas, decreasing = T)
betas = betas[ind]
rt = rt[,ind]
syb = syb[ind]
betas
```

```{r}
rt.A = rt[,1:9]
rt.B = rt[,10:18]
```

b\)

The portfolio without short selling has N+2 constraints with:

$$
\text{Amtx}=\begin{bmatrix} \mu & 1 & I_n\end{bmatrix}, \text{and }\begin{bmatrix}m \\ 1 \\ 0 \end{bmatrix} 
$$

Where meq = 2.

The minimum variance portfolio removes the first constraint. So we have:

```{r}
library(quadprog)
N = 9
ones = rep(1,9)
zeros = rep(0,9)
rt.A.mu = apply(rt.A, 2, mean)
rt.A.S = var(rt.A)

Amat = cbind(ones, diag(N))
bvec = c(1, zeros)
out = solve.QP(Dmat = rt.A.S, dvec = zeros, Amat = Amat, bvec = bvec, meq = 1)
w.min = out$solution; w.min = w.min*(abs(w.min) > 10e-7);names(w.min) = syb[1:9];
mu.min = sum(w.min*rt.A.mu);sd.min = sqrt(2*out$val)
w.min
c(return = mu.min, risk = sd.min)
```

c\)

We obtain a portfolio on the tangent line with any pre-specified return m then we re=scale so w sums to 1.

```{r}
rt.B.mu = apply(rt.B, 2, mean)
rt.B.S = var(rt.B)
mu.f = 3.31/52
Amat = as.matrix(rt.B.mu - mu.f)
bvec = 0.3
out = solve.QP(Dmat = rt.B.S, dvec = zeros, Amat = Amat, bvec = bvec, meq = 1)
w.star = out$solution; names(w.star) = syb[10:18]
w.T = w.star/sum(w.star)
cat("Portfolio:"); w.T

mu.T = sum(w.T*rt.B.mu);s.T = sqrt(2*out$value)/sum(out$solution)
c(return = mu.T, risk = s.T)

```

d\)

We use that $w = (.5,.5)$ then expected return is $w^T\mu$ and the variance covariance matrix is $w \Sigma w^T$.

```{r}
w = c(.5,.5)
mean_est <- sum(w*c(mu.min,mu.T))
Sigma = diag(c(sd.min, s.T)^2)
Sigma[1,2] = Sigma [2,1] = as.vector( t(w.min)%*%cov(rt.A,rt.B)%*%w.T)
c(mean = mean_est)
Sigma
```

e\)

The portfolio with the highest Sharpe ratio is the Tangency Porfolio. For the sake of simplicity I will say that short selling is not allowed.

```{r}
port.mu = c(mu.min, mu.T)
port.s = Sigma
mu.f = 3.31/52
Amat = as.matrix(port.mu - mu.f)
bvec = 0.3

zeros = rep(0,2)
out = solve.QP(Dmat = port.s, dvec = zeros, Amat = Amat, bvec = bvec, meq = 1)
w.star2 = out$solution; names(w.star) = c("A","B")
w.T2 = w.star2/sum(w.star2)
cat("Portfolio:"); w.T2

mu.T2 = sum(w.T2*port.mu);s.T2 = sqrt(2*out$value)/sum(out$solution)
c(return = mu.T2, risk = s.T2)
```

f\)

```{r}
weights = c(w.T2[1]*w.min,w.T2[2]*w.T)
weights
```

So we have test statistic:

$$
\frac{m-k+1}{k}y^TS^{-1}y \sim F_{k,m-k+1}
$$

In this case we fit model first:

```{r}
Mkt_excess <- Mkt - Rf
rt_excess <- apply(rt, 2, function(u) u - Rf)
fit_f <- lm(rt_excess ~ Mkt_excess)
fit_f
```

```{r}
alpha = fit_f$coefficients[1,]
n = dim(rt)[1]
N = dim(rt)[2]
y = t(weights)%*%alpha
m11 = sum_fits[[1]]$cov.unscaled[1,1]
Sig = 1/n*t(resid(fit))%*%resid(fit)
S = n*m11*t(weights)%*%Sig%*%weights
k = 18
stat_f <- (n - k - 1)/k*y%*%solve(S)%*%y
stat_f
c(stat = stat_f, p.value = 1 - pf(stat_f, df1 = k, df2 = n - k - 1 ))
```

We cannot reject the null hypothesis of CAPM.

g\)

```{r}
return_g <- mu.f*.15 + .85*mu.T2
risk_g <- .85*s.T2
c(return = return_g, risk = risk_g)
```
