---
title: "STAT 631 Lab 1"
author: "Jack Cunningham (jgavc@tamu.edu)"
date: 08/27/2024
date-format: short
format:
  pdf:
    include-in-header:
      - text: |
          \usepackage{amsmath}
editor: visual
engine: knitr
---

```{r}
dat = read.csv("Stock_bond.csv", header = TRUE)
head(dat)
tail(dat)
```

```{r}
names(dat)
attach(dat)
```

```{r}
par(mfrow = c(1, 2))
plot(GM_AC)
plot(F_AC)
```

```{r}
plot(GM_AC, type = "l")
```

```{r}
plot(GM_AC, type = "b")
```

```{r}
n = dim(dat)[1]
GMReturn = GM_AC[-1]/GM_AC[-n] - 1
FReturn = F_AC[-1]/ F_AC[-n] - 1
par(mfrow = c(1, 1))
plot(GMReturn, FReturn)
```

Problem 1:

Do the GM and Ford returns seem positively correlated? Do you notice any outlying returns? If "yes," do outlying GM returns seem to occur with outlying Ford returns?

Yes GM and Ford returns appear to be positively correlated. Outlying returns for GM also seem to occur with outlying Ford returns as well.

Problem 2:

Compute the log returns for GM and plot the returns versus the log returns. How highly correlated are the two types of returns? (The R function cor() computes correlations.)

```{r}
GM_returns <- diff(GM_AC)
GM_log_returns <- diff(log(GM_AC))
plot(GM_returns, GM_log_returns)
```

```{r}
cor(GM_returns, GM_log_returns)
```

The two returns appear to be very highly correlated, with a correlation $\rho=0.9137$.

Problem 3

```{r}
#Computting Returns
MSFTReturn = MSFT_AC[-1] / MSFT_AC[-n] - 1
MRKReturn = MRK_AC[-1] / MRK_AC[-n] - 1
```

```{r}
par(mfrow = c(1, 2))

plot(MSFTReturn, MRKReturn)
```

```{r}
cor(MSFTReturn, MRKReturn)
```

The two returns visually appear to not be highly correlated. Computing correlation shows a slight positive correlation. Outlying returns seem to occur at the same time for both somewhat frequently but not always.

2.4.2 Simulations

```{r}
niter = 1e5
below = rep(0, niter)
set.seed(2009)

for (i in 1:niter)
{
  r = rnorm(45, mean = 0.05/253,
            sd = 0.23/sqrt(253))
  logPrice = log(1e6) + cumsum(r)
  minlogP = min(logPrice) # minimum price over next 45 days
  below[i] = as.numeric(minlogP < log(950000))
}
mean(below)
```

Problem 4 What is the probability that the value of the stock will be below \$950,000 at the close of at least one of the next 45 trading days? To answer this question, run the code below.

Using our simulation we get an estimate of 0.50988 as the probability of the stock being below \$950,000 in one of the next 45 trading days.

```{r}
niter = 1e5
final_stock_value = rep(0, niter)
day_return_earned = rep(0, niter)
set.seed(2009)

for (i in 1:niter)
{
  r = rnorm(100, mean = 0.05/253,
            sd = 0.23/sqrt(253))
  logPrice = log(1e6) + cumsum(r)
  upper = ifelse(logPrice > log(1100000), yes = 1, no = 0)
  lower = ifelse(logPrice < log(950000), yes = 1, no = 0)
  if(sum(upper) > 1){
  upper_day = min(which(upper == 1))
  }
  if(sum(lower) > 1){
  lower_day = min(which(lower == 1))
  }
  
  if(sum(upper) == 0 & sum(lower) == 0){
    final_stock_value[i] = logPrice[length(logPrice)]
    day_return_earned[i] = 100
  } else if(sum(upper) > 0 & sum(lower) == 0){
    final_stock_value[i] = logPrice[upper_day]
    day_return_earned[i] = upper_day
  } else if(sum(upper) == 0 & sum(lower) > 0){
    final_stock_value[i] = logPrice[lower_day]
    day_return_earned[i] = lower_day
  } else{
    which_first = upper_day < lower_day
    if(which_first){
    final_stock_value[i] = logPrice[upper_day]
    day_return_earned[i] = upper_day
    } else{
    final_stock_value[i] = logPrice[lower_day]
    day_return_earned[i] = lower_day
    }
  }
}
```

Problem 5 What is the probability that the hedge fund will make a profit of at least \$100,000?

```{r}
large_profit_chance <- (sum(final_stock_value > log(1100000)))/niter
large_profit_chance
```

There is a 37.2% chance that the hedge fund will make a profit of at least \$100,000.

Problem 6 What is the probability the hedge fund will suffer a loss?

```{r}
loss_probability <- (sum(final_stock_value < log(1000000)))/niter
loss_probability
```

The loss probability of this strategy is 58.3%.

Problem 7 What is the expected profit from this trading strategy?

```{r}
expected_profit <- sum(final_stock_value - log(1000000))/niter
exp(expected_profit)
```

The expected profit of this strategy is 1 dollar.

Problem 8

What is the expected return? When answering this question, remember that only \$50,000 was invested. Also, the units of return are time, e.g., one can express a return as a daily return or a weekly return. Therefore, one must keep track of how long the hedge fund holds its position before selling.

```{r}
expected_daily_return <- (sum((final_stock_value - log(1000000))/day_return_earned))
exp(expected_daily_return)
exp(expected_daily_return)*30
```
