---
title: "STAT 631 Risk"
author: "Jack Cunningham (jgavc@tamu.edu)"
date: 11/08/2024
date-format: short
format:
  pdf:
    include-in-header:
      - text: |
          \usepackage{amsmath}
editor: visual
engine: knitr
---

```{r}
load("Risk.RData")
```

1\)

We use the profile likelihood method to fit the multivariate t distribution.

```{r}
library(MASS)
library(mnormt)

df = seq(1,8, 0.01)
loglik_p = c()
for(i in 1:length(df)){
  fit = cov.trob(y4 , nu = df[i])
  loglik_p[i] = sum(dmt(y4, mean = fit$center, S = fit$cov, df = df[i], log = T))
}

nu = df[which.max(loglik_p)]
cat("The MLE of degrees of freedom:", paste(nu))
```

Our estimates are:

```{r}
est = cov.trob(y4, nu = nu, cor = T)
names(est)
```

The MLE for the mean vector is:

```{r}
est$center
```

The MLE of the scale matrix Lambda is:

```{r}
est$cov
```

The MLE of Covariance is:

```{r}
est$cov*nu/(nu - 2)
```

b\)

The tangency portfolio allowing short selling has the following explicit solution:

$$
w_t=\frac{\Sigma^{-1}\mu_{ex}}{1^T\mu_{ex}}
$$

Where $\Sigma$ is the covariance matrix:

```{r}
y.4.S = est$cov*nu/(nu - 2)
y.4.S
```

And mu.ex is the excess return, taking the MLE of the mean vector and subtracting by the risk free rate:

```{r}
mu.f = 3.5/52
m.ex = est$center - mu.f
ones = rep(1,4)
```

We can now calculate $w_t$:

```{r}
IS = solve(y.4.S)
aT = as.numeric((t(ones)%*%IS%*%m.ex))
w4.T = 1/aT*(IS%*%m.ex)
mu4.T = as.numeric(t(w4.T)%*%est$center)
s4.T = sqrt(as.numeric(t(w4.T)%*%y.4.S%*%w4.T))
cat("Tangency Portfolio for y4:"); w4.T[,1];
cat("Portfolio return is:", mu4.T, "\t with risk", s4.T)

```

c)  

If we choose to have 20% of assets on the risk free asset and 80% on risky assets we choose a point on the tangency line.

```{r}
w4.c = c(w4.T*.8,.2)
names(w4.c) = c(syb4, "RF")
w4.c
```

Then we can compute the expected return and risk easily, through:

$$
\mu_p=\mu_f+w(\hat{\mu_T}-\mu_f),\hat{\sigma}_p=w\hat{\sigma}_T
$$

```{r}
mu.4.c = .8*mu4.T + .2*mu.f
S.4.c = .8*s4.T
cat("Portfolio for y4, 20% in RF:"); w4.c;
cat("Portfolio return is:", mu.4.c, "\t with risk", S.4.c)
```

We use the following to find the distribution of the portfolio:

$$
w^TY\sim t_{\nu}(w^T\mu,w^T\Lambda w)
$$

We need to take the risk from our last step, square it and transition it to the scale parameter:

```{r}
scale.c = sqrt(S.4.c^2 * (nu - 2)/nu)
```

So the distribution of the portfolio is: $t_{\hat{\nu}}(.2415,1.311287)$ where $\hat{\nu}=$ 4.45.

d\)

Using this distribution we can calculate the one-week VaR and ES with S = 50000.

```{r}
S = 50000
alpha = c(.05,0.01)
q.t = qt(alpha, df = nu); VaR.t = -S*(mu.4.c + scale.c*q.t);
ES.t = S*(-mu.4.c + scale.c*dt(q.t, nu)/alpha*(nu+q.t^2)/(nu-1))
VaR.t = VaR.t/100; ES.t = ES.t/100
output <- rbind(alpha, VaR.t, ES.t)
output
```

2\)

First we fit a t-distribution for each return series:

```{r}
n = dim(y8)[1];N = dim(y8)[2]
nu = c(); mu = c(); lambda = c()
for(i in 1:N){
  start = list(m = mean(y8[, i]), s = sd(y8[,i]), df = 4)
  fit = fitdistr(y8[, i ], "t", start, lower = c(-100, 0.0001, 0.1))
  nu[i] = fit$est["df"]
  mu[i] = fit$est["m"]
  lambda[i] = fit$est["s"]
}
stat = cbind(nu, mu, lambda)
colnames(stat) = c("DF", "Mu", "Lambda")
rownames(stat) = syb8
stat
```

```{r}
alpha = 0.05
S = 50000
q.t = qt(alpha,df = nu); VaR.t = -S*(mu +lambda*q.t);
ES.t = S*(-mu + lambda*dt(q.t,nu)/alpha*(nu*q.t^2)/(nu-1))
VaR.t = VaR.t/100;ES.t = ES.t/100
stat_2 <- cbind(VaR.t, ES.t)
colnames(stat_2) <- c("VaR.t 0.05", "ES.t 0.05")
rownames(stat_2) <- syb8
stat_2

```

b\)

We use the explicit solution below again:

$$
w_t=\frac{\Sigma^{-1}\mu_{ex}}{1^T\mu_{ex}}
$$
The means and covariance matrix of y8 are below:

```{r}
y.mu = apply(y8,2,mean);y.mu
y.S = var(y8); y.S
```

We create m.ex by subtracting the risk free rate, which is 3.5/52 in this case. We then we the explicit solution for $w_t$ from before:

```{r}
m.ex = y.mu - mu.f
ones <- rep(1,N)
IS = solve(y.S)
aT = as.numeric((t(ones)%*%IS%*%m.ex))
w8.T = 1/aT*(IS%*%m.ex)
w8.T
```
