---
title: "STAT 631 Homework 6"
author: "Jack Cunningham (jgavc@tamu.edu)"
date: 10/14/2024
date-format: short
format:
  pdf:
    include-in-header:
      - text: |
          \usepackage{amsmath}
editor: visual
engine: knitr
---

```{r}
load("HW05.Rdata")
```

3\)

Computing the mean and variances of the return data:

```{r}
y.mu = apply(y,2,mean);y.mu
y.S = var(y);y.S
```

Defining various vectors for convenience:

```{r}
N <- dim(y)[2]
ones <- rep(1,N)
zeros <- rep(0,N)
```

Given the constraints $-0.2 \leq w_i \leq0.4$ we can set up our optimization for the solveLP function per the below:

$$
\text{cvec}=\begin{bmatrix}  \mu \\ -\mu\end{bmatrix},\text{Amat}=\begin{bmatrix}I_n & 0 \\ 0 & I_n \\ 1^T & -1^T \end{bmatrix},\text{bvec}=\begin{bmatrix} (0.4)1 \\ (0.2)1 \\ 1\end{bmatrix}
$$

First we obtain the feasible range of returns, by minimizing and maximizing our linear programming problem:

```{r}
## Feasible range of m.R with b1 = 0.4 and b2 = 0.2
library(linprog)
b1 = 0.4; b2 = 0.2
cvec = c(y.mu, -y.mu)
Amat.lp = rbind(diag(2*N), c(ones, -ones))
bvec.lp = c(rep(b1,N), rep(b2,N),1)
inequ = c(rep("<=", 2*N), "=")
min.lp = solveLP(cvec = cvec, bvec = bvec.lp, Amat = Amat.lp,
                 lpSolve=T, const.dir = inequ, maximum = F)
max.lp = solveLP(cvec = cvec, bvec = bvec.lp, Amat = Amat.lp,
                 lpSolve=T, const.dir = inequ, maximum = T)
mu.lim = c(lower = min.lp$opt, upper = max.lp$opt); mu.lim;
```

Since in part c we need to plot the efficient frontier we find efficient portfolios for the whole range of feasible returns:

```{r}
library(quadprog)
m.R = seq(round(mu.lim[1] + .0005, 3), round(mu.lim[2]-.0005, 3), 0.001)
sd.R = c();
wmat = matrix(nrow = length(m.R), ncol = N); colnames(wmat) = syb;
Amat = cbind(y.mu, ones, -diag(N), diag(N));
for(i in 1:length(m.R)){
  bvec = c(m.R[i],1, rep(-b1, N), rep(-b2,N))
  out = solve.QP(y.S, dvec = zeros, Amat=Amat, bvec = bvec, meq = 2);
  sd.R[i] = sqrt(2*out$value)
  wmat[i,] = out$solution
}
```

a\)

In our previous step we found all the portfolios on the efficient frontier. The minimum variance is the portfolio where risk is minimized, we find this portfolio below:

```{r}
i.min = which.min(sd.R); ## index with smallest sd
w.min = wmat[i.min,];mu.min = m.R[i.min]; sd.min = sd.R[i.min];
cat("Minimum variance portfolio:"); w.min
c(return = mu.min, risk = sd.min)
```

b\)

The tangency portfolio is the portfolio on the efficient frontier that maximizes the Sharpe ratio $\frac{E(R_p) - \mu_\text{f}}{\sigma_{R_p}}$. We find this portfolio below:

```{r}
mu.f = 4.37/52

i.T = which.max((m.R - mu.f)/sd.R);
w.T = wmat[i.T,]; mu.T = m.R[i.T]; sd.T = sd.R[i.T]
cat("Tangency portfolio:"); w.T
c(return = mu.T, risk = sd.T)
```

c\)

```{r}
s.min = sd.min
s.T = sd.T
y.sd = sqrt(diag(y.S))

plot(sd.R, m.R, type = "l", xlim = c(0,max(sd.R)), xlab = "Risk", ylab = "Return")
lines(sd.R[m.R > mu.min], m.R[m.R > mu.min], lwd = 2.5, col = "red3");  ## efficient frontier
lines(c(0,s.T), c(mu.f, mu.T), col = "navy")
points(0,mu.f, pch = 19, col = "royalblue") ## risk-free asset return
points(s.T,mu.T, pch = 21, col ="royalblue", bg = "green", cex = 1.2) ## tangency portfolio
points(s.min, mu.min, pch = 19, col = "red3") ## minimum variance portfolio
for(i in 1:N){ ## individual stock
	text(y.sd[i], y.mu[i], syb[i], font = 2)
}

legend("topleft",c("Efficient Frontier", "Efficient Portfolio","Tangency Portfolio"),lty=c(1,1, NA),lwd=c(4,3, 3),pch=c(NA,NA, 21), 
col=c("red3","royalblue","royalblue"), pt.bg=c("","","green"), pt.cex = c(NA,NA, 1.5), y.intersp = 1.2,bty = "n", xjust = 5)

```

d\)

We are looking for an efficient portfolio with allowed risk of 2.5%. The risk of the tangency portfolio is `r sd.T`%, since this is higher than our allowed risk we will have a portfolio with weight $w_t$ in the tangency portfolio and $w_f$ in the risk free asset with $w_t + w_f =1$.

We find the appropriate weights by using the following formula for the risk of our desired portfolio, $\sigma_p$:

$$
\sigma_p=w_t\sigma_t
$$

Since we know the allowed risk of 2.5% and the risk of the tangency portfolio we can solve for $w_t$:

$$
w_t=\frac{\sigma_p}{\sigma_t}
$$

From this we can find the weight in the risk free asset through:

$$
w_f=1-w_t
$$

And the expected return of the portfolio, $E(R_p)$ , through:

$$
E(R_p)=w_f \mu_f+w_t \mu_t
$$

We compute each below:

```{r}
sd_p_d <- 2.5
w_t_d <- sd_p_d/sd.T
w_f_d <- 1 - w_t_d
mu_p_d <- w_f_d*mu.f + w_t_d*mu.T
w_p_d <- c(w_f_d, w_t_d*w.T);names(w_p_d) = c("Risk Free", syb)
cat("Portfolio:");w_p_d
c(return = mu_p_d, risk = sd_p_d)
```

e\)

We are looking for an efficient portfolio with a target return of 0.55%. The expected return of the tangency portfolio is `r mu.T`%, since the target return is greater than what is achieved by the tangency portfolio we need to find a portfolio on the efficient frontier with an expected return of 0.55%. Earlier we computed efficient portfolios for varying target returns, including 0.55%. It is below:

```{r}
i.e <- which(m.R == 0.55)
w.e <- wmat[i.e,]
mu.e <- 0.55
sd.e <- sd.R[i.e]
cat("Portfolio:");w.e
c(return = mu.e, risk = sd.e)
```

f\)

We are looking for an efficient portfolio with target return of 0.85%. Recall that the feasible range of expected returns given our constraints is (`r min.lp$opt` , `r max.lp$opt`). Since the target return of 0.85% lies outside that range there is no efficient portfolio for this target return without loosening our constraints.
