---
title: "STAT 631 Homework 2"
author: "Jack Cunningham (jgavc@tamu.edu)"
date: 09/09/2024
date-format: short
format:
  pdf:
    include-in-header:
      - text: |
          \usepackage{amsmath}
editor: visual
engine: knitr
---

## 

1\)

We have function $f^\star(y|\xi)$ below:

$$
f^\star(y|\xi)=\begin{cases} f(y\xi) & y<0 \\ f(y/\xi) & y \geq0\end{cases}
$$

We are tasked with integrating $f^{\star}(y|\xi)$:

$$
\int_{-\infty}^{\infty}f^\star(y|\xi)=\int_{-\infty}^0f(y\xi)dy+\int_0^\infty f(y/\xi)dy
$$

Let $u=y\xi$ so $du=\xi dy$. Let $w=y/\xi$ so $dw=\frac{1}{\xi}dy$. Then we have:

$$
\xi\int_{-\infty}^0f(u)du+\xi^{-1}\int_{0}^{\infty}f(w)dw
$$

$$
\xi(F(0)-F(-\infty))+\xi^{-1}(F(\infty)-F(0))
$$

Since $f$ is a symmetric pdf about 0 $F(0)=P(y<0)=1/2$. Also $F(-\infty)=0$ and $F(\infty)=1$ from basic rules about CDFs. Then we have our desired result:

$$
\xi(1/2-0)+\xi^{-1}(1-1/2)=\frac{1}{2}(\xi+\xi^{-1})
$$

2\)

a\)

The formula of Kurtosis for a random variable with a finite fourth moment is its fourth standardized moment.

$$
Kur=\frac{E[(x-\mu)^4]}{\sigma^4}
$$

We know that $Kur=3$ for a normal distribution. This implies that $E[(x-\mu)^4]=3\sigma^4$ for the normal distribution.

We have the below discrete density mixture:

$$
f(x)=.95f_1(x)+.05f_2(x)
$$

Where $f_1(x)$ is the density function of $N(0,1)$ and $f_2(x)=N(0,10)$.

A helpful fact for this problem is that $\mu^{(k)}=\sum_{i=1}^np_iE_{f_i}[x_i^{k}]$.Essentially when we are computing a moment we can compute each distribution's moment, multiply them by their weight, and sum them together.

This leads to $f(x)$ having $\sigma^2=p_1\sigma_1^2+p_2\sigma_2^2=.9(1)+.1(10)=1.9$. And $E[(x-\mu)^4]=p_1E_{f_1}[(x-\mu_1)^4]+p_2Ef_2[(x-\mu_2)^4]$. So using what we know about the fourth central moment of the normal distribution we have:

$$
E[(x-\mu)^4]=p_1(3\sigma_1^4)+p_2(3\sigma_2^4)
$$

Then we can compute kurtosis with the below:

$$
Kur=\frac{3(p_1\sigma_1^4+p_2\sigma_2^4)}{(p_1\sigma_1^2+p_2\sigma_2^2)^2}
$$

```{r}
p_1 <- .9
sigma_1 <- 1
p_2 <- .1
sigma_2 <- sqrt(10)

Kur <- 3*(p_1*sigma_1^4+p_2*sigma_2^4)/(p_1*sigma_1^2+p_2*sigma_2^2)^2
Kur
```

The kurtosis for this discrete mixture distribution is `r round(Kur,3)`.

b\)

Building off what I worked on in the previous question we have the below after substituting $\sigma_1=1,\sigma_2=\sigma,p_1=p,p_2=(1-p_1)$.

$$
Kur(p,\sigma)=\frac{3(p+(1-p)\sigma^4)}{(p+(1-p)\sigma^2)^2}
$$

3\)

```{r message=FALSE}
library(quantmod)
```

```{r}
getSymbols("^GSPC",from = "2005-01-01", to = "2024-08-01")
x = weeklyReturn(Ad(GSPC),type = "log")*100
n = dim(x)[1]
```

a\)

```{r message=FALSE}
library(rugarch)
```

```{r}
dists = c("std", "sstd", "ged", "sged", "nig", "jsu")
fits = vector("list", 6)
for (i in 1:6) fits[[i]]  = fitdist(dists[i], x)
```

b\)

```{r fig.height= 10, fig.width = 10}
den = density(x, adjust = 0.75)
x0 = den$x; y0 = den$y
par(mfrow = c(2,3))

for(i in 1:length(dists)) {
  plot(x0, y0, type = "l", main = dists[i], ylim = c(0,0.3), xlab = "returns",
  ylab= "density")
  est = fits[[i]]$pars
  yi = ddist(dists[i], x0, mu = est["mu"], sigma = est["sigma"], skew =
  est["skew"], shape = est["shape"])
lines(x0, yi, col = "red3")
}

```

From this first plot we can see that the GED and skewed GED both seem to peak sharper than the kernel density estimate. The GED estimate also appears to struggle in the right shoulder region. The other four estimates fit well in the middle region. Next let's take a look at the left tail region.

```{r fig.height= 10, fig.width = 10}
par(mfrow = c(2,3))

for(i in 1:length(dists)) {
  plot(x0, y0, type = "l", main = dists[i], ylim = c(0,0.03), xlim = c(-22,-8), xlab = "returns",
  ylab= "density")
  est = fits[[i]]$pars
  yi = ddist(dists[i], x0, mu = est["mu"], sigma = est["sigma"], skew =
  est["skew"], shape = est["shape"])
lines(x0, yi, col = "red3")
}
```

Looking at the extreme tail values we can get an idea of how quickly each distributions tail dissipates. We can see that the GED distribution appears to zero out around $x=13$, too early compared to the kernel density estimator. The skewed GED distribution seems to taper a bit later around $x=15$, still too early. The standardized t distribution does well here, it it intercepting around the middle of the kernel density estimator and tapers off completely at around $x=19$. The other three distributions do well but seem to slightly overestimate the density in the $x=8$ to $x=11$ range.
