---
title: "STAT 631 Homework 8"
author: "Jack Cunningham (jgavc@tamu.edu)"
date: 11/01/2024
date-format: short
format:
  pdf:
    include-in-header:
      - text: |
          \usepackage{amsmath}
editor: visual
engine: knitr
---

```{r}
source("FM_Functions.R")
source("Factor_Tests.R")
load("HW08.RData")
attach(FF5)
```

1\)

The Fama-French 3 factor model is the below:

$$
Y_t=\alpha+B^TF_t+\epsilon_t, \quad E[\epsilon_t|F_t]=0,\quad E[\epsilon_t \epsilon_t^T|F_t]=\Sigma_{\epsilon}
$$

Where $F=\begin{bmatrix}\text{Excess Return Market Portfolio} & \text{Small Minus Big} & \text{High Minus Low}\end{bmatrix}$, these are each economic vectors with length n.

```{r}
Yt = apply(Rt,2, function(x) x-RF); dimnames(Yt)[[2]] = syb;
n = dim(Yt)[1]; N = dim(Yt)[2]; p = 3
fit = lm(Yt ~ Mkt.RF + SMB + HML); sfit = summary(fit)
```

a\)

```{r}
betas = coef(fit)[-1,]
R.Squared = c(); for(i in 1:N) R.Squared[i] = sfit[[i]]$r.squared
coef.plot(R.Squared, coef(fit)[-1,],labs = syb)
```

From the R squared plot we see that the three factor Fama French model performance varies greatly. Let's take a look at the breakdown by industry:

```{r}
table(Hi_R.Sq = R.Squared > 0.5, by_industry)
```

Generally R-Squared isn't very high for these assets. There are only two that exceed 0.5, both are in the technology industry Microsoft and Autodesk.

```{r}
table(Hi_R.Sq = R.Squared < 0.2, by_industry)
```

R-Squared is particularly low for the Food industry, six of the seven stocks have an R-Squared beneath 0.2. This indicates that the three factor Fama French model does not perform well for this industry.

```{r}
table(Aggressive = coef(fit)[2,] > 1, by_industry)
```

On an industry level we see that that Food and Heath Care are not aggressive compared to market returns while Technology generally is. Entertainment is more of a mixed bag.

```{r}
library(tidyverse)
compare <- data.frame(
  Stock = syb,
  Beta = coef(fit)[2,],
  Industry = by_industry
)
compare |> 
  group_by(Industry) |> 
  summarise(Average_Beta = mean(Beta))
```

By taking a look at the average Beta we can see that the Food Industry has a Beta of 0.4132 on average. Healthcare, despite being not aggressive compared to the market, is far closer to 1 in comparison.

b\)

To identify the individual assets that don't follow the FF-3-factor model we use the t-test for $H_0:\alpha_i=0$ that is automatically computed from the lm function.

```{r}
Alpha = c()
for(i in 1:N){
  Alpha = rbind(Alpha, sfit[[i]]$coef[1, ])
}
dimnames(Alpha)[[1]] = syb
Alpha_df <- data.frame(Alpha, Industry = by_industry)
Alpha_df |> 
  filter(Pr...t.. < .05)
```

There are four individual assets that do not follow the FF-3 factor model, Live Nation Entertainment, Paramount, Warner Brothers Discovery and Pediatric Medical Group. The first three are in the entertainment industry and the last is in healthcare.

c\)

We are testing the hypothesis that $H_0:\alpha = 0$. If we reject this hypothesis this indicates that the FF-3 factor does not hold for all 27 assets. We perform the Wald and Likelihood Ratio Tests.

```{r}
alpha <- coef(fit)[1, ]
res = resid(fit); Sig.e = 1/n*t(res)%*%res
m11 = sfit[[1]]$cov.unscaled[1,1]
var.alpha = m11*Sig.e
p = 3

wald.fun(est = alpha, est.var = var.alpha, n =  n, p = p)
res.0 = resid(lm(Yt~Mkt.RF + SMB + HML - 1))
Sig.e0 = 1/n*t(res.0)%*%res.0
lrt.fun(sig = Sig.e, sig0 = Sig.e0,n = n)
```

Both the Wald and Likelihood test ratios have a similar result with p value $\approx.271$. We cannot reject the null hypothesis that the FF-3 factor holds for all 27 assets.

d\)

```{r}
wald = c(); lrt = c()
for(i in industry){
  ind = which(by_industry == i)
  wald = rbind(wald, wald.fun(alpha[ind], m11*Sig.e[ind,ind],n = n, p = p))
  lrt = rbind(lrt, lrt.fun(Sig.e[ind,ind], Sig.e0[ind,ind], n = n))
}

rownames(wald) = rownames(lrt) = industry
cat("Wald test by industry:"); wald
cat("LRT by industry:"); lrt
```

All industries cannot reject the null hypothesis that the FF-3 factor model holds for their respective stocks at a significance level of 0.05. However there is still a significant difference between the industries. At a significance level of 0.1 both entertainment and healthcare would reject the null hypothesis. The Food industry however has a p.value $\approx 0.96$, the evidence strongly suggests the FF-3 factor model holds well for this industry.

e\)

The sample covariance approach has $N(N+1)/2$ estimates. With $N = 27$ this is 378 estimates.

The model based approach has $(p+1)(N+p/2)$ estimates. With $N=27,p=3$ this is 114 estimates.

```{r}
resid.summary(res)
```

We can see high correlation between stocks in the same industry but small correlation between stocks in different industries. This indicates that our assumption of a diagonal covariance matrix could be unreasonable. This would call into question any inference we obtain from our model. We should formally check for this and consider an industry factor model as a possible option.

f\)

The test for block-diagonal matrices tests $H_0:\Sigma=\text{diag}\{\Sigma_{11},\dots,\Sigma_{kk}\}$ has test statistic:

$$
\text{LRT}=-\log\frac{\det(\hat{\Sigma})}{\det(\hat{\Sigma}_{11})\dots \det(\hat{\Sigma}_{kk})}
$$

This statistic is approximately $\chi^2_v$, with degrees of freedom $v=\frac{1}{2}(d^2-\sum_{i=1}^kd_i^2)$.

```{r}
cov.diag.test(Sig.e, Ns = Ns, n = n, p = p)
```

This test rejects the null hypothesis of block-diagonal matrices.

We also test whether the full matrix is diagonal. This is an adaption of the previous test, we have $d_i=1,i=1,...,d$. Then the statistic is $-\log \det(\hat{\text{Corr(y))}}$, with degrees of freedom $v=\frac{1}{2}d(d-1)$.

```{r}
cov.diag.test(Sig.e, Ns = rep(1,N), n = n, p = p)
```

This test rejects the null hypothesis of a diagonal covariance matrix.

2\)

```{r}
fa.none = factanal(Yt,3,rotation = "none")
print(fa.none)
```

The first factor has all positive coefficients and are relatively similar, it seems to be a shared market component.

The second factor has negative, with the exception of a near zero coefficient for Danone SA, coefficient for all stocks in the food industry. This appears to be an industry factor.

The third factor looks like an aggressiveness component. It is positive for all tech and entertainment stocks with generally negative values for health and food stocks. This comports with our previous analysis of betas for each stock.

b\)

```{r}
p = 3
Zt = apply(Yt, 2, function(u) (u-mean(u))/sd(u))
fa = factanal(Zt, p, scores = "Bartlett", rotation = "none")
B = t(fa$loading)
Ft.fa = fa$scores
resid_mat = Zt - Ft.fa %*% B
resid.summary(resid_mat)
```

There is still correlation but it is less grouped by industry. It is less severe between particular stocks, before in the FFA-3-F model correlation was particularly strong for certain pairs. This model makes the assumption of a diagonal covariance matrix more reasonable.
